RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R

### 2015-01-08 manuscripts/manuscript.Rmd
add section explicitly showing validation

@@

The returned list does not include the data elements to which the
metadata are attached. Therefore, a different approach, documented in
the package vignette, is recommended for accessing the metadata attached
to data elements.

@@

Because NeXML expresses all metadata using the RDF standard, and
stores them compliant with RDFa, they can be extracted as an RDF
graph, queried, analyzed, and mashed up with other RDF data, local or
on the web, using a wealth of off-the-shelf tools for working with RDF
(see @W3C_2014 or @Hartig_2012). Examples for these possibilities are
included in the `RNeXML` vignette, and the package also comes with a
demonstration that can be run from R using the following command:
`demo("sparql", "RNeXML")`).

@@

Writing out metadata annotations as in the above proposal may seem
tedious. However, by virtue of `RNeXML`'s API it is straightforward to
define functions that encapsulate these details. As a demonstration,
`RNeXML` comes with functions `simmap_to_nexml()` and `nexml_to_simmap()`
that map bidirectionally between NeXML files with `simmap` metadata
according to the above proposal, and the extension of the `ape::phylo`
class devised by @Revell_2012 in the R package `phytools`. The `RNeXML`
package vignette illustrates how this allows seamlessly generating and
reading NeXML files containing `simmap` data, and using the various
functions from `phytools` designed for `simmap` objects [@Revell_2012].

>>>>

## Validating NeXML

File validation is a central feature of the NeXML format which ensures
that any properly implemented NeXML parser will always be able to read
the NeXML file. The function takes the path to any NeXML file and returns 
`TRUE` to indicate a valid file, or `FALSE` otherwise, along with a
display of any error messages generated by the validator.

```{r }
nexml_validate("geospiza.xml")
```

The `nexml_validate()` function performs this validation
using the online NeXML validator (when a network connection is available),
which performs additional checks not expressed in the NeXML schema itself [@Vos_2012].
If a network connection is not available, the function falls back on the 
schema validation method from the `XML` package [@XML].

@@

The returned list does not include the data elements to which the
metadata are attached. Therefore, a different approach, documented in
the metadata vignette, is recommended for accessing the metadata attached
to data elements.

@@

Because NeXML expresses all metadata using the RDF standard, and
stores them compliant with RDFa, they can be extracted as an RDF
graph, queried, analyzed, and mashed up with other RDF data, local or
on the web, using a wealth of off-the-shelf tools for working with RDF
(see @W3C_2014 or @Hartig_2012). Examples for these possibilities are
included in the `RNeXML` SPARQL vignette, and the package also comes with a
demonstration that can be run from R using the following command:
`demo("sparql", "RNeXML")`).

@@

Writing out metadata annotations as in the above proposal may seem
tedious. However, by virtue of `RNeXML`'s API it is straightforward to
define functions that encapsulate these details. As a demonstration,
`RNeXML` comes with functions `simmap_to_nexml()` and `nexml_to_simmap()`
that map bidirectionally between NeXML files with `simmap` metadata
according to the above proposal, and the extension of the `ape::phylo`
class devised by @Revell_2012 in the R package `phytools`. The `RNeXML`
simmap vignette illustrates how this allows seamlessly generating and
reading NeXML files containing `simmap` data, and using the various
functions from `phytools` designed for `simmap` objects [@Revell_2012].
